---
title: "new engines in maize"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{new engines in maize}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(maize)
```

### kernels for the ksvm engine

{maize} is continuing down the path of least known engines, novelties, and unique ML methods. 

Initially, {maize} was created to add additional kernel support to {parsnip}, outside of the known linear, polynomial, and radial basis function kernels. This added a handful of very odd kernels from within the {kernlab} library. After this first leap, there were a few customized kernels added and registered within {kernlab} and {maize}, based on other efforts by Python and Julia open sourced packages.    

From there, the package started cooking up more stuff, like adding a few {recipes} related to kernel principal component analysis and Hebbian algorithms, which added new **step_** functions to the package. 


### beyond the ksvm engine

While the package is built around supporting kernels and, well, support vectors, there are many other engines that can be added to {maize}. A lot of these are within the {kernlab} library, given we have already wrapped the "ksvm", the next step would be adding some of the other engines within {kernlab}. (Another handful of engines come from {mildsvm} - shoutout to @jrosell for finding these. support for these will be tested in the future). 

#### RVM

#### KQR

#### KKNN


