---
title: "perchance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{perchance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


### Post-Processing in {maize}???

**based entirely on the {probably} package implementation and conformal inference scripts. **

{probably} contains a int_conformal_quantile method for prediction intervals. The underlying method in probably uses a random forest quantile regression. Within {maize}, int_conformal_quantile_svm has been added. There was a simple swap for a quantile regression method with SVMs. (This method uses {qrsvm} which is not on CRAN.)

```{r}
library(qrsvm)
library(dplyr)
library(parsnip)
library(maize)
library(recipes)
library(workflows)
library(kernlab)
library(probably)
```

### point and interval regressions
```{r}

set.seed(31415)

  # note: qrsvm does not work for non-numeric values
corn_df <- corn_data |> mutate(type = as.numeric(factor(type, levels = levels(type))))

corn_train <- corn_df |> dplyr::sample_frac(.80)
corn_cal   <- corn_df |> dplyr::anti_join(corn_train) |> head(30)
corn_test  <- corn_df |> dplyr::anti_join(corn_train) |> tail(30)


# SVM base model 
svm_spec <-
  svm_laplace() |>
  set_mode("regression")

svm_wflow <-
  workflow() |>
  add_model(svm_spec) |>
  add_formula(height ~ .)

svm_fit <- fit(svm_wflow, data = corn_train)


# probably's implementation:
rf_int <- probably::int_conformal_quantile(svm_fit, corn_train, corn_cal,
  level = 0.80
)

# experimental QRSVM interface for CQR: 
svm_int <- int_conformal_quantile_svm(svm_fit, corn_train, corn_cal,
  level = 0.80, cost = 1000, degree = 5
)

svm_int

```

### predict with intervals
```{r}
cqr_rf  <- predict(rf_int,  corn_test)
cqr_svm <- predict(svm_int, corn_test)


# renaming pre-plotting
names(cqr_rf)  <- c(".pred", ".pred_lo_rf",  ".pred_hi_rf")
names(cqr_svm) <- c(".pred", ".pred_lo_svm", ".pred_hi_svm")
```

### check conformal quantile prediction intervals

## SVM with QRF & QRSVM Prediction Intervals

```{r}
library(ggplot2)

corn_test |>
  dplyr::bind_cols(cqr_svm) |>
  dplyr::bind_cols(cqr_rf |> dplyr::select(-.pred)) |>
  ggplot() +
  geom_ribbon(aes(x = kernel_size, ymin = .pred_lo_svm, ymax = .pred_hi_svm), 
              fill = "midnightblue", alpha = .5) +
  geom_ribbon(aes(x = kernel_size, ymin = .pred_lo_rf,  ymax = .pred_hi_rf), 
              fill = "orange", alpha = .5) +
  geom_point(aes(x = kernel_size, y = height, color = "truth")) +
  geom_point(aes(x = kernel_size, y = .pred,  color = "pred")) + 
  theme_minimal() +
  labs(subtitle = "Corn Height Prediction Intervals")
  
```

## SVM with Linear & SVM Calibration 


```{r}
calibration_df <-  predict(svm_fit, corn_cal) |> cbind(corn_cal)


# probably's implementation:
lin_cal_fit <- calibration_df |>
            probably::cal_estimate_linear(height, smooth = TRUE)

# experimental svm calibration (vanilladot or polydot)
svm_cal_fit <- calibration_df |>
            maize::cal_estimate_svm(height, smooth = TRUE)

# preds
cal_corn_test <- 
predict(svm_fit, corn_test) |>
  cbind(corn_test)
  
# calibrated preds:
maize::cal_apply_regression(svm_cal_fit, cal_corn_test)

```

