---
title: "perchance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{perchance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


### Post-Processing in {maize}???

**based entirely on the {probably} package implementation for conformal inference & calibration. Hence the pun perchance? **

The {probably} package implements the int_conformal_quantile method for generating prediction intervals. This method utilizes random forest quantile regression as its underlying algorithm. In the {maize} package, a new method called int_conformal_quantile_svm has been introduced. This implementation replaces the original quantile regression random forest {quantregForest} approach with a quantile regression SVM {qrsvm}. It's worth noting that the int_conformal_quantile_svm method relies on the {qrsvm} package, which is currently not available on CRAN.

#### Key Components
{probably} Package    
Method: int_conformal_quantile   
Algorithm: `quantregForest::quantregForest()`   
     
{maize} Package       
Method: int_conformal_quantile_svm      
Algorithm: `qrsvm::qrsvm()`     

#### Implementation Details
The int_conformal_quantile_svm method in {maize} represents a straightforward substitution of the RF algorithm. Instead of using random forests, it employs SVMs for generating prediction intervals. This modification allows users to leverage the potential benefits of SVMs in certain prediction scenarios.

#### Dependency Note
The {qrsvm} package, which is required for the SVM-based implementation in {maize}, is not currently hosted on CRAN. Users interested in utilizing the int_conformal_quantile_svm method should be aware of this external dependency and may need to install it from github.com/frankiethull/qrsvm.

```{r, message=FALSE, warning=FALSE}
library(qrsvm)
library(dplyr)
library(parsnip)
library(maize)
library(recipes)
library(workflows)
library(kernlab)
library(probably)
```

### point and interval regressions

The following demonstration begins with an initial point forecast using a Support Vector Machine (SVM) with a Laplacian kernel. Subsequently, we utilize our out-of-bag (OOB) sample to construct conformal prediction intervals. In this comparison, we will evaluate both the int_conformal_quantile method from the {probably} package and the int_conformal_quantile_svm method from the {maize} package. This analysis will provide insights into the performance and characteristics of these two approaches for generating prediction intervals.
```{r, message=FALSE, warning=FALSE}

set.seed(31415)

  # note: qrsvm does not work for non-numeric values
corn_df <- corn_data |> mutate(type = as.numeric(factor(type, levels = levels(type))))

corn_train <- corn_df |> dplyr::sample_frac(.80)
corn_cal   <- corn_df |> dplyr::anti_join(corn_train) |> head(30)
corn_test  <- corn_df |> dplyr::anti_join(corn_train) |> tail(30)


# SVM base model 
svm_spec <-
  svm_laplace() |>
  set_mode("regression")

svm_wflow <-
  workflow() |>
  add_model(svm_spec) |>
  add_formula(height ~ .)

svm_fit <- fit(svm_wflow, data = corn_train)


# probably's implementation:
rf_int <- probably::int_conformal_quantile(svm_fit, corn_train, corn_cal,
  level = 0.80
)

# experimental QRSVM interface for CQR: 
svm_int <- int_conformal_quantile_svm(svm_fit, corn_train, corn_cal,
  level = 0.80, cost = 1000, degree = 5
)

svm_int

```

### predict with intervals
```{r}
cqr_rf  <- predict(rf_int,  corn_test)
cqr_svm <- predict(svm_int, corn_test)


# renaming pre-plotting
names(cqr_rf)  <- c(".pred", ".pred_lo_rf",  ".pred_hi_rf")
names(cqr_svm) <- c(".pred", ".pred_lo_svm", ".pred_hi_svm")
```

### check conformal quantile prediction intervals

## SVM with QRF & QRSVM Prediction Intervals

```{r}
library(ggplot2)

corn_test |>
  dplyr::bind_cols(cqr_svm) |>
  dplyr::bind_cols(cqr_rf |> dplyr::select(-.pred)) |>
  ggplot() +
  geom_ribbon(aes(x = kernel_size, ymin = .pred_lo_svm, ymax = .pred_hi_svm, fill = "QR-SVM"), alpha = .5) +
  geom_ribbon(aes(x = kernel_size, ymin = .pred_lo_rf,  ymax = .pred_hi_rf,  fill = "QR-RF"),  alpha = .5) +
  geom_point(aes(x = kernel_size, y = height, color = "truth")) +
  geom_point(aes(x = kernel_size, y = .pred,  color = "pred")) + 
  theme_minimal() +
  labs(subtitle = "Corn Height Prediction Intervals")
  
```

## SVM with Linear & SVM Calibration 

In addition to conformal prediction intervals, the {maize} package offers a post-process calibration engine that is similar to the one in {probably}, but utilizes Support Vector Machines (SVMs). The {probably} package includes the cal_estimate_linear function, which calibrates point forecasts using a post-model to address bias in the initial fit. {maize} provides a comparable function called cal_estimate_svm. This function follows a similar framework to cal_estimate_linear, but employs either a vanilla or polynomial kernel, depending on the user's selection in the "smooth" argument. The following section demonstrates the application of this calibration approach.

#### Key Components
{probably} Package    
Method: cal_estimate_linear   
Algorithms:   
- `stats::glm()` is used when smooth is set to `FALSE`   
- `mgcv::gam()` is used when smooth is set to `TRUE`  
     
{maize} Package       
Method: cal_estimate_svm      
Algorithms:        
- `kernlab::ksvm(kernel = "vanilladot")` is used when smooth is set to `FALSE`    
- `kernlab::ksvm(kernel = "polydot")` is used when smooth is set to `TRUE`   

#### Implementation Details
The cal_estimate_svm method in {maize} represents a straightforward substitution of the linear calibration. Instead of using linear models, it employs SVMs for post-calibration. This modification allows users to leverage the potential benefits of SVMs in certain prediction scenarios.



```{r}
calibration_df <-  predict(svm_fit, corn_cal) |> cbind(corn_cal)


# probably's implementation:
lin_cal_fit <- calibration_df |>
            probably::cal_estimate_linear(height, smooth = TRUE)

# experimental svm calibration (vanilladot or polydot)
svm_cal_fit <- calibration_df |>
            maize::cal_estimate_svm(height, smooth = TRUE)

# preds
cal_corn_test <- 
predict(svm_fit, corn_test) |>
  cbind(corn_test)
  
# calibrated preds:
maize::cal_apply_regression(svm_cal_fit, cal_corn_test) |> head(10)

```

